Simulation of an Attic Fire in a Wood Frame Residential Structure - Chicago, IL

### 2015-05-01 Studies/2012_Chicago/Report/response.tex

\documentclass[12pt]{article}
\usepackage[pdftex]{graphicx}
\frenchspacing
\begin{document}

\section{Comments from the Editor}
\begin{itemize}
\item We would appreciate more efforts in assessing or estimation somehow the error or uncertainty of your results. \\
{\it A section was added to reflect uncertainties in FDS.}
\item An effort to read some of the work published in Fire Technology on the topics at hand would be appreciated by the editorial board. \\
{\it Added addtional references where appropriate.}
\item There are too many figures (20); please consider removing some or creating an Appendix where the 5 least important figures are presented. \\
{\it Moved figures}
\item Please consider to submit the input file/s for this work as Electronic Supplementary Material. \\
{\it Will do}
\end{itemize}

\section{Comments from Reviewer 1}
\begin{itemize}
\item It may be better to summarize the account in the author's own words, and to link it more closely to Table 1. This would be particularly helpful for section 3.4, since the ventilation conditions are not explained in great detail. \\
{\it The authors were able to view portions of video taken during the incident that allowed for the estimation of failure times and failure areas of exterior walls and windows. This is addressed in the text in Section 3.4. Since this case deals with the death of a firefighter the authors decided to leave the decription of events to narrative from the official NIOSH report.}
\item I miss, however, a more critical point of view that would be essential in guiding fire modellers as to what the limits of applicability of these models is. In this context I would require that the authors elaborate on how they approached the modelling of under-ventilated conditions. FDS is known for running into problems when flaming in vitiated air occurs. In figure 7 the authors present external flaming out of the rear window. Is this realistic? FDS tends to present combustion at vents, even when the flames in the interior have already been extinguished due to lack of oxygen. Is the flow of hot gases into the hallway after the door breakage accompanied by flaming combustion? \\
{\it The authors have witnessed this type of attic vent fire behavior in studies that are currently underway (cf. Fig.~\ref{fig:eave}) and previous work conducted by Underwriters' Laboratory. Therefore, they belive this behavoir to be realistic.}

\begin{figure}[!ht]
\centering
\includegraphics[width=.6\textwidth]{../Figures/PA020340}
\caption{External buirning throught attic eave}
\label{fig:eave}
\end{figure}

\item The authors discard the secondary simulation based on the similarity of the overall HRR, but an alternative ventilation pattern could potentially alter the flow path inside the building. They should explain in more detail why this is not the case in this scenario (or provide images of the flow path and temperature distributions of the secondary simulation). \\
{\it Text was added to state that in both cases the flow path path was the same and the hazard to the firefighters were similar.}
\item I further think that the beginning of section 5.2 is misleading. Surely second-degree burn injuries are only produced by direct contact with objects at 55C. Air at 55C would not burn the skin. It would be useful to relate the temperatures (gas and wall) to heat fluxes, as this would be a more correct approach in terms of human perception and injuries. \\
{\it The section has been updated.}
\end{itemize}

@@

\item On the numerical resolution examination [see p. 8, lines 22-27], it would be helpful to explain that higher numbers equate to better resolution as the non-user may not appreciate this and think that being outside the range makes it an unreliable calculation. \\
{\it Text added to clarify.}
\item HRR: The graphic [Fig. 6, p.10] is fine and accompanying text is also good. One thing that is not clear, but may be helpful to the reader is to explain that some of that combustion is occurring outside the structure [See Fig. 7, p. 12]. This is important given the discussion of temperatures, etc. If the authors know the amount of HRR within the structure, that may be an interesting/helpful value for the readers. \\
{\it Addressed by other reviewer's comments. See section in text.}

@@

\item Fig. 14, p. 16: This photo shows the door and discusses the crease, which I can see. I assume from my reading that this is the interior face of the door and that the top of the door is towards the background. This should be clarified in the caption. \\
{\it Clarified in the text.}
\item Fig. 20, p. 24 \& Fig. 21, p. 25: These graphics are good overall. Unfortunately, the cross-hatching is difficult to interpret in some places. Furthermore, it will be impossible to read this if it is reproduced in black and white. \\
{\it Images were removed to address excess figures but a reference was made to report where figures live.}

### 2015-05-01 Studies/2012_Chicago/Report/firetech_chicago.tex
Chicago_Fire: update report and response file for FireTech, add attic fire images, and cleanup input file.

The fire, which is described in Section~\ref{fire}, started in the attic and spread to the rear porch. Discussion of the fire's impact on the interior conditions is provided in Section~\ref{results}. Figure~\ref{fig:alpha_ex} shows the exterior of the front and rear of the structure after the incident. Note that the majority of the damage occurred in the attic and rear porch on the second floor. Fully dimensioned drawings of the interior of the first and second floors are shown in Appendix~\ref{app_draw} in Figs.~\ref{fig:first_floor} and \ref{fig:second_floor}, respectively.

@@

@@

In Fig.~\ref{fig:hrr}, the solid line represents the baseline FDS simulation (this follows the noted door openings from the incident). The dashed line represents an alternative series of events in which the rear door does not get opened and the subsequent breach of the rear porch wall on the left side of the structure does not occur. The dash-dotted line represents the prescribed (or nominal) HRR that was input into FDS. As shown in Fig.~\ref{fig:hrr}, the HRR output from FDS is less than the prescribed HRR because the fire in the structure becomes ventilation limited. The maximum HRR value is slightly less than 9~MW before the fire becomes ventilation limited and begins to decay. The HRR is reduced to 5~MW as the oxygen inside the structure is consumed. There is an insufficient amount of oxygen in the structure to burn all of the fuel, therefore combustion occurs locally at the vents where fresh air is available. Figure~\ref{fig:smv_ext_fire} shows flaming combustion through the failed window on the rear side of the structure and attic breach just after the second fire ignites at 40~s. After additional wall breaches occur at 45~s and 55~s (see Table~\ref{tab:vents}), the abrupt change in ventilation causes the HRR to increase to 9~MW before it decreases to approximately 6~MW. The HRR then increases towards a sustainable steady-state value.

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=.675\textwidth]{../Figures/smv_exterior_fire}
\caption{Smokeview rendering of exterior combustion occurring on the rear side of the structure.}
\label{fig:smv_ext_fire}
\end{figure}

@@

\subsection{Assessing the Hazard}
\label{hazard}
A person is susceptible to second-degree burn injuries if skin temperatures exceed 55~$^{\circ}$C (130~$^{\circ}$F)~\cite{contactburn}. Although firefighters wear protective gear, gear only offers a finite amount of protection. The polycarbonate material in the facepiece of a self-contained breathing apparatus begins to soften when the material temperature reaches approximately 140~$^{\circ}$C (284~$^{\circ}$F)~\cite{mensch2011emergency}. Structural fire fighting coats and pants are tested to withstand temperatures up to 260~$^{\circ}$C (500~$^{\circ}$F) or 10~kW/m$^2$~\cite{nfpa2013standard} for less than 2.5~min. Exposure to elevated temperatures can result in a significant amount of heat transferred to the firefighter, putting him or her at risk.

@@

\section{Dimensioned Drawings}
\label{app_draw}

\begin{figure}[!ht]
\centering
\includegraphics[width=1\textwidth]{../Figures/50th_Place_1st_Floor}
\caption[Dimensioned drawing of first floor.]{Dimensioned drawing of first floor. The measurements are accurate to within 15~cm (6~in).}
\label{fig:first_floor}
\end{figure}

\begin{figure}[!ht]
\centering
\includegraphics[width=1\textwidth]{../Figures/50th_Place_2nd_Floor}
\caption[Dimensioned drawing of second floor.]{Dimensioned drawing of second floor. The measurements are accurate to within 15~cm (6~in).}
\label{fig:second_floor}
\end{figure}

>>>>

The fire, which is described in Section~\ref{fire}, started in the attic and spread to the rear porch. Discussion of the fire's impact on the interior conditions is provided in Section~\ref{results}. Figure~\ref{fig:alpha_ex} shows the exterior of the front and rear of the structure after the incident. Note that the majority of the damage occurred in the attic and rear porch on the second floor. Fully dimensioned drawings of the interior of the first and second floors can be found in~\cite{Weinschenk:Chicago}.

@@

\subsection{Model Uncertainy}
\label{mod_uncertain}
FDS can provide valuable insight into how a fire may have developed. However, the model is only a simulation. The model output is dependent on a variety of input values such as material properties, time lines, geometry, and ventilation openings. Since complete knowledge of every detail of the fire site, fuel load, or fire timeline is never available; estimations are incorporated into the model. For example, the estimation of the energy release rate of an initial ``source fire'' as a starting point for fire development and spread throughout the structure is a necessary part of recreating this fire scenario. These estimations and others used in this simulation are further described in Section~\ref{model} of this report.

The ability of the FDS model to predict accurately the temperature and velocity of fire gases has been previously evaluated by conducting experiments, both lab-scale and fullscale, and measuring quantities of interest. For relatively simple fire driven flows, such as buoyant plumes and flows through doorways, FDS predictions are within the experimental uncertainty of the values measured in the experiments~\cite{FDS_Validation_Guide}. For example, if a gas flow velocity is measured at 0.5 m/s (1.6 ft/s) with an experimental uncertainty of $\pm$ 0.05 m/s ($\pm$ 0.2 ft/s), the FDS model gas flow velocity predictions were also in the range between 0.45 m/s and 0.55 m/s (1.5 ft/s and 1.8 ft/s). In large scale fire tests reported in \cite{FDS_Validation_Guide}, FDS temperature predictions were found to be within 15 \% of the measured temperatures and the FDS heat release rates were predicted to within 20 \% of the measured values. Therefore, the results should be considered as ranges to address these uncertainties.

@@

In Fig.~\ref{fig:hrr}, the solid line represents the baseline FDS simulation (this follows the noted door openings from the incident). The dashed line represents an alternative series of events in which the rear door does not get opened and the subsequent breach of the rear porch wall on the left side of the structure does not occur. The dash-dotted line represents the prescribed (or nominal) HRR that was input into FDS. As shown in Fig.~\ref{fig:hrr}, the HRR output from FDS is less than the prescribed HRR because the fire in the structure becomes ventilation limited. The maximum HRR value is slightly less than 9~MW before the fire becomes ventilation limited and begins to decay. The HRR is reduced to 5~MW as the oxygen inside the structure is consumed. There is an insufficient amount of oxygen in the structure to burn all of the fuel, therefore combustion occurs locally at the vents where fresh air is available. Figure~\ref{fig:smv_ext_fire} shows flaming combustion through the failed window on the rear side of the structure and attic breach just after the second fire ignites at 40~s. Figure~\ref{fig:smv_ext_fire} also shows a photograph taken during an attic fire experiment that cofirms this type of fire behavoir. After additional wall breaches occur at 45~s and 55~s (see Table~\ref{tab:vents}), the abrupt change in ventilation causes the HRR to increase to 9~MW before it decreases to approximately 6~MW. The HRR then increases towards a sustainable steady-state value.

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=.5\textwidth]{../Figures/smv_exterior_fire}
\includegraphics[width=.45\textwidth]{../Figures/PA020413}
\caption{Smokeview rendering of exterior combustion occurring on the rear side of the structure (left) and photograph from ongoing attic fire experiments (right).}
\label{fig:smv_ext_fire}
\end{figure}

@@

\subsection{Assessing the Hazard}
\label{hazard}
A person is susceptible to second-degree burn injuries if skin temperatures exceed 55~$^{\circ}$C (130~$^{\circ}$F)~\cite{contactburn}. Although firefighters wear protective gear, gear only offers a finite amount of protection. The polycarbonate material in the facepiece of a self-contained breathing apparatus begins to soften when the material temperature reaches approximately 140~$^{\circ}$C (284~$^{\circ}$F)~\cite{mensch2011emergency}. Structural fire fighting coats and pants are tested to withstand temperatures up to 260~$^{\circ}$C (500~$^{\circ}$F) or 10~kW/m$^2$~\cite{nfpa2013standard} for less than 2.5~min. Exposure to elevated temperatures can result in a significant amount of heat transferred to the firefighter, putting him or her at risk. In this case, according the NIOSH report the victim was not wearing gloves or flash hood~\cite{NIOSH:Bowyer}.

@@
